---
title: "Untitled"
author: "FENDAWN F. RECENTES"
date: "12/2/2022"
output:
  pdf_document: 
    fig_width: 5
    fig_height: 4
  html_document: default
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Helper packages and Modeling packages
```{r}
# Helper packages
library(dplyr)       # for data manipulation
library(ggplot2)     # for data visualization
library(stringr)     # for string functionality
library(gridExtra)   # for manipulating the grid

# Modeling packages
library(tidyverse)  # data manipulation
library(cluster)     # for general clustering algorithms
library(factoextra)  # for visualizing cluster results
```
\

#### Loading of Data and Removing of Any Missing Values
```{r}
data("iris")

#To remove any missing value that might be present in the data, type this:
df <- na.omit(iris)
```
\

#### Scaling/standardizing the data
```{r}
#we start by scaling/standardizing the data
df <- scale(df[c(1:4)])
head(df)
```
\

#### Start 2 clusters
```{r}
#start at 2 clusters
k2 <- kmeans(df, centers = 2, nstart = 25)
str(k2)
```
\

#### Plot the 2 clusters
```{r}
#plot the 2 clusters
fviz_cluster(k2, data = df)
```
\

#### Each clusters data
```{r}
#get the each clusters data
df %>%
  as_tibble() %>%
  mutate(cluster = k2$cluster,
         Species = row.names(iris)) %>%
  ggplot(aes(Sepal.Length, Sepal.Width, color = factor(cluster), label = Species)) +
  geom_text()

k3 <- kmeans(df, centers = 3, nstart = 25)
k4 <- kmeans(df, centers = 4, nstart = 25)
k5 <- kmeans(df, centers = 5, nstart = 25)
```
\

#### Plots to compare
```{r}
# plots to compare
p1 <- fviz_cluster(k2, geom = "point", data = df) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, geom = "point",  data = df) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, geom = "point",  data = df) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, geom = "point",  data = df) + ggtitle("k = 5")

grid.arrange(p1, p2, p3, p4, nrow = 2)
```
\

#### Determining Optimal Number of Clusters
```{r}
#Determining Optimal Number of Clusters
set.seed(123)
```
\

#### Function to compute total within-cluster sum of square 
```{r}
# Function to compute total within-cluster sum of square 
wss <- function(k) {
  kmeans(df, k, nstart = 10 )$tot.withinss
}
```
\

#### Compute and plot wss for k = 1 to k = 15
```{r}
# Compute and plot wss for k = 1 to k = 15
k.values <- 1:15
```
\

#### Extract wss for 2-15 clusters
```{r}
# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```
\

#### Or use this
```{r}
#or use this
fviz_nbclust(df, kmeans, method = "silhouette")
```
\

#### Compute gap statistic
```{r}
# compute gap statistic
set.seed(123)
gap_stat <- clusGap(df, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
```
\

#### Print the result
```{r}
# Print the result
print(gap_stat, method = "firstmax")

```
\

#### 
```{r}
fviz_gap_stat(gap_stat)
```
\

#### Compute k-means clustering with k = 2
```{r}
# Compute k-means clustering with k = 2
set.seed(123)
final <- kmeans(df, 2, nstart = 25)
print(final)
```
\

#### Final data
```{r}
# Final data
fviz_cluster(final, data = df)
```





